---
title: "How to create R packages"
output: html_notebook
---

This is a tutorial on a few steps to create R packages.

## 0) Check if the name you chose to your package is available

```{r, eval = F}
library(available)

available("howtoRpackage", browse = F)
```


## 1) Create the R package within R

We can create the structure of a package using the function `create_package` from the `usethis` package. This create the very minimal structure before building the package.

```{r}
library(usethis)

# create the package structure
usethis::create_package("howtoRpackage")
```

## 2) Define a license

One can define a license by hand or using a few pre-determined functions from the `usethis` package. If we use the function, this will modify both the `DESCRIPTION` file and also add a  `LICENSE.md` file.

```{r}
# Define GPL3 as our license
usethis::use_gpl3_license(name = "Bernardo Niebuhr, Elie Gurarie et al")
```

## 3) Add some functions

Now we add some functions. We can create them by hand (just copy and paste now) within the `R` folder. Now we can build

## 4) Add packages upon which the functions depend

```{r}
# Necessary packages
usethis::use_package("sp", type = "Imports")

# Suggested packages
usethis::use_package("plyr", type = "Suggests")
```

## 5) Add data - now we add both raw and ready to use

First we create the `raw-data` folder using the `use_data_raw` function. It will also create a script for preparation of the raw data into the R data format, `.rda`.

```{r}
# add raw data
usethis::use_data_raw()
```

Now, we put the our dataset in the `data-raw` folder. Then we can edit the script for preparing the dataset. There we read and prepare the data, then we save it as `.rda` using, e.g. the function `usethis::use_data` or `save()`:

```{r}
# read data
juveniles <- read.csv("data-raw/subsetGPSlocs.csv")

# create data for use in the Rpackage
usethis::use_data(juveniles, overwrite = TRUE)

# or
dir.create("data/")
save(juveniles, file = "data/juveniles.rda")
```

## 6) Use the package

Now we can use the package, and then save this as an example script on how
to use our functions.

## 7) Add some `roxyxen2` documentation to the script

We can add some info to the functions, using the `roxygen2` notation, and it will be transformed into documentation in the `man` folder.

```{r}
usethis::use_git(message = "initial commit")
```

